Hazelnut: A Bidirectionally Typed \\ Structure Editor Calculus

### 2016-07-06 hazelnut-popl17.tex
reformatted sum type extension as two figures: the main figure, and movement
finished 3.1 except for a few minor todos
tweaked colors
tweaks
sec 3.2 done
writing about hazelnut

\def\OPTIONConf{0}%
%\documentclass[12pt]{article}
\documentclass[preprint,9pt]{sigplanconf}
% The following \documentclass options may be useful:
%
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% authoryear    To obtain author/year citation style instead of numeric.
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{ stmaryrd }
\usepackage{mathpartir}
\usepackage{stmaryrd}
\usepackage{wasysym}
\usepackage{extarrows}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\definecolor{light-gray}{gray}{0.9}
\usepackage{mathpazo}
\usepackage{colortab}
\usepackage{url}
\usepackage{todonotes}
\usepackage{listings}
\lstset{tabsize=2,
basicstyle=\ttfamily\fontsize{8pt}{1em}\selectfont}
\usepackage{microtype}
\sloppy
\def \TirNameStyle #1{\small\rulename{#1}}
\renewcommand{\MathparLineskip}{\lineskiplimit=.3\baselineskip\lineskip=.3\baselineskip plus .2\baselineskip}

@@

\usepackage[compact]{titlesec}
\titlespacing*{\section}{0pt}{4pt}{2pt}
\titlespacing*{\subsection}{0pt}{4pt}{2pt}
\titlespacing*{\subsubsection}{0pt}{4pt}{2pt}
\titlespacing*{\paragraph}{0pt}{4pt}{2pt}
\setlength{\skip\footins}{3pt}

@@

Hazelnut is a foundational type-theoretic account of typed structure editing, rather than an end-user tool itself. To that end, we describe how Hazelnut's rich metatheory, which we have mechanized in Agda, guides the definition of an extension to the calculus. We also discuss various plausible evaluation strategies for terms with holes, and in so doing reveal connections with gradual typing and contextual modal type theory (the Curry-Howard interpretation of contextual modal logic.) Finally, we  discuss how Hazelnut's semantics lends itself to implementation as a functional reactive program. Our reference implementation is written using \lstinline{js_of_ocaml}.

@@

\section{Extending Hazelnut}\label{sec:extending}
\todo{write this section} Add sum types.

H-expression

Z-expression + Erasure rules

Statics

Movement actions

Construct actions

\input{sumtypes}

@@

>>>>

\def\OPTIONConf{1}%
%\documentclass[12pt]{article}
\documentclass[preprint,9pt]{sigplanconf}
% The following \documentclass options may be useful:
%
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% authoryear    To obtain author/year citation style instead of numeric.
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{ stmaryrd }
\usepackage{mathpartir}
\usepackage{stmaryrd}
\usepackage{wasysym}
\usepackage{extarrows}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\definecolor{light-gray}{gray}{0.9}
\usepackage{soul}
\setulcolor{red}
\usepackage{mathpazo}
\usepackage{colortab}
\usepackage{url}
\usepackage{todonotes}
\usepackage{listings}
\lstset{tabsize=2,
basicstyle=\ttfamily\fontsize{8pt}{1em}\selectfont}
\usepackage{microtype}
\sloppy
\def \TirNameStyle #1{\small\rulename{#1}}
\renewcommand{\MathparLineskip}{\lineskiplimit=.3\baselineskip\lineskip=.3\baselineskip plus .2\baselineskip}

@@

\usepackage[compact]{titlesec}
\titlespacing*{\section}{0pt}{4pt}{2pt}
\titlespacing*{\subsection}{0pt}{4pt}{2pt}
\titlespacing*{\subsubsection}{0pt}{4pt}{2pt}
\titlespacing*{\paragraph}{0pt}{4pt}{2pt}
\setlength{\skip\footins}{3pt plus 1px minus 5px}

@@

\todo{rephrase}Hazelnut is a foundational type-theoretic account of typed structure editing, rather than an end-user tool itself. To that end, we describe how Hazelnut's rich metatheory, which we have mechanized in Agda, guides the definition of an extension to the calculus. We also discuss various plausible evaluation strategies for terms with holes, and in so doing reveal connections with gradual typing and contextual modal type theory (the Curry-Howard interpretation of contextual modal logic.) Finally, we  discuss how Hazelnut's semantics lends itself to implementation as a functional reactive program. Our reference implementation is written using \lstinline{js_of_ocaml}.

@@

\section{Extending Hazelnut}\label{sec:extending}
%% \todo{write this section} Add sum types.
%% H-expression
%% Z-expression + Erasure rules
%% Statics
%% Movement actions
%% Construct actions
\input{sumtypes}

@@

\paragraph{Gradual typing: Has Matched Relation}
%
A. Rastogi, A. Chaudhuri, and B. Hosmer. The ins and outs of gradual type inference.
\\
Definition 3.1 is similar to our ``has matched'' relations:
\\
\url{https://www.cs.umd.edu/~avik/papers/iogti.pdf}

This paper defines ``has matched'' relation for arrow, just like us:
\\
\url{http://cimini.info/publications/Gradualizer_Draft.pdf}
\\
They credit this technique to the following paper:
\\
\url{http://www.cs.ubc.ca/~rxg/ptsgp.pdf}
\\
But, the Rastogi paper came earlier.
